---
title: "Data Visulization"
author: "Jordan Wozniak"
date: "9/18/2025"
categories: [welcome, fantasyteam, blog2]
image: ""
format: html
---

# A.I. Disclosure {#sec-aiDisclosure}

AI Disclosure: AI was not used.

Link to AI Transcript: not applicable

# Preamble {#sec-preamble}

## Install Libraries
```{r}
#install.packages("remotes")
#remotes::install_github("DevPsyLab/petersenlab")
#remotes::install_github("FantasyFootballAnalytics/ffanalytics")

#update.packages(ask = FALSE)
```

## Load Libraries
```{r}
#| message: false
#| warning: false

library("petersenlab")
library("XICOR")
library("psych")
library("DescTools")
library("psych")
library("correlation")
library("effectsize")
library("GGally")
library("ggridges")
library("tidyverse")
```

## Load Data

```{r}
load("player_stats_weekly.RData")
```

# Abstract {#sec-abstract}

This study examined the relationship between age and fantasy football performance using weekly NFL data. Visualizations created in R showed only a weak link between age and fantasy points, with wide variation across players. Results suggest that role and opportunity are stronger predictors of fantasy production than age alone.

# Background {#sec-background}

## Research Questions {#researchQuestion} 

Does age matter when it comes to fantasy performance?

## Hypothesis {#sec-hyptothesis}

I believe that age does play a factor into a player's fantasy performance. The older they get, the less they get.

## Justification {#sec-justification}

You have dominate players like Aaron Rodgers, Tom Brady, and Matthew Stafford, who in their "hay-days," where fantasy point machines. During their last seasons in the league, around or over the age of 40, have significantly slowed down. There are some outliers and the difference of who is on that team at the time. Is there a correct assumption to say that the older the player is, the worse they get, or is it irrelevant?

## Method {#sec-method}

The method that I want to use here is to use a table with all players and their seasonal fantasy point numbers, compared to their age, all in a correlation plot. This is better for this, in my opinion, is that you get the average fantasy point numbers by the x and y that is chosen, in this case it would be age and season fantasy points. This is also cleaner and produces an average, which is what we want to create a correct answer. What we are trying to find here is that there is a positive skew, leaning towards the younger age versus the older age.

#Results {#sec-results}
We are going to load player stats weekly, where it will look at all the fantasy points each week for a player, with the added information of their age, and compare.

```{r}
ggplot2::ggplot(
  data = player_stats_weekly %>%
    filter(position %in% c("QB", "RB", "WR", "TE")),
  aes(x = age, y = fantasyPoints)
) +
  geom_point() +
  geom_smooth(method = "lm", color = "red") +
  geom_smooth(method = "gam", color = "blue") +
  labs(
    x = "Age",
    y = "Fantasy Points",
    title = "Correlation between Age and Fantasy Points"
  )
```

# Discussion {#sec-discussion}

From running the data, my original hypothesis was somewhat wrong, let me explain. First and foremost, I think it is okay to have a wrong hypothesis, if you run the correct data and correctly reflect on the findings. The data shown is weekly fantasy performances on the side, where the age is at the bottom. There are a lot of players that start off very young, and staying the league doesn't really show a downgrade in fantasy. However, this could relate to actual performance instead of fantasy performance, but that's not what we were specifically looking for.

If I were to look at some of the outliers, they start to go down in a downward, where they are lest likely to have an "insane" week. Could it be said that it is safer to pick a younger player because there is a lower range, sure.

# Strengths and Limitations {#sec-strengthsLimits}

There are always strengths and weaknesses when it comes to using data, example used in the NFL, to draw different conclusions on many events. A strength is that we, as a class, were able to use the data provided that listed plentiful of stats that the NFL documents. We can compare many different statistics to others and draw comparisons on why this event happened.

However with the limitation, in my own study, age doesn't always predict performance, in this case, fantasy points. There could be other factors, like injuries, position on the depth chart, relations with the team, position within the field, and other contributing factors.

# Conlusion {#sec-conclusion}

This analysis examined the relationship between age and fantasy football production. The results suggest that there is little to no strong correlation between age and fantasy points, with most players producing within a wide range regardless of age. While the trend lines indicate a slight upward slope, the variation across individuals is much larger than any age-based pattern.

# References {#sec-references}

https://bigguyfantasysports.com/wide-receiver-age-cliff-in-fantasy-is-it-real/#:~:text=There%20are%20arbitrary%20ages%20for,is%20thought%20to%20be%2030.

# Session Info {#sec-sessionInfo}
```{r}
sessionInfo()
```