---
title: ""
author: "Jordan Wozniak"
date: "10/23/2025"
categories: []
image: ""
format: html
---

# Preamble {#sec-preamble}

## Install Libraries
```{r}
#install.packages("remotes")
#remotes::install_github("DevPsyLab/petersenlab")
#remotes::install_github("FantasyFootballAnalytics/ffanalytics")

#update.packages(ask = FALSE)
```

## Load Libraries
```{r}
#| message: false
#| warning: false

library("petersenlab")
library("tidyverse")
```

## Load Data

```{r}
#| eval: false

load(file = "../../data/player_stats_weekly.RData")
load(file = "../../data/players_projections_weekly.RData")
load(file = "../../data/nfl_playerIDs.RData")
```

```{r}
options(scipen = 999) # prevent scientific notation
```

# A.I. Disclosure {#sec-aiDisclosure}

AI Disclosure: AI was not used.

Link to AI Transcript: not applicable


# Abstract {#sec-abstract}

**Background:**

**Method:**

**Results:**

**Discussion:**

# Background {#sec-background}

I will evaluate whether a wide receiver’s prior-game receiving yards predict hitting ≥60 yards in the next game, using binary-classification accuracy

## Research Question

For NFL wide receivers, how accurately and how well-calibrated does prior-game receiving yards predict a next-game outcome of ≥60 receiving yards, and is this signal strong enough to guide start/sit decisions?

## Hypothesis

Prior-game receiving yards will best predict a WR reaching ≥60 yards next game; specifically, higher prior yards increase the odds of repeating the 60+ threshold with decent calibration.

## Justification

60 yards is a sustainable number for receivers to achieve. Making it an accurate number to see. Through analysis in fantasy leagues, having a wide receiver that achieves this number on a consistent basis can predict that they are a reliable receiver to draft/keep.

# Method {#sec-method}



# Data {#sec-data}
```{r}
library(dplyr)
library(petersenlab)

player_stats_weekly_example <- player_stats_weekly %>%
  filter(position == "WR") %>%
  arrange(player_id, season, week) %>%
  group_by(player_id, season) %>%
  mutate(
    receiving_yards_priorGame = dplyr::lag(receiving_yards),
    yds60_next = as.integer(dplyr::lead(receiving_yards >= 60))
  ) %>%
  ungroup() %>%
  filter(!is.na(receiving_yards_priorGame), !is.na(yds60_next))

player_stats_weekly_example %>%
  select(player_id, season, week, receiving_yards, receiving_yards_priorGame, yds60_next) %>%
  head()
```

```{r}
petersenlab::accuracyAtCutoff(
  predicted = player_stats_weekly_example$receiving_yards_priorGame,
  actual    = player_stats_weekly_example$yds60_next,
  cutoff    = 59 # 1 before 60, to see how many fit
)
```

# Results {#sec-results}

# Discussion {#sec-discussion}

# Strengths and Limitations {#sec-strengthsAndLimitations}


# Citation {#sec-Citation}

https://www.rotoviz.com/2016/08/using-air-yards-to-better-predict-receiving-yards/

# Session Info {#sec-sessionInfo}
```{r}
sessionInfo()
```
