---
title: ""
author: "Jordan Wozniak"
date: "10/23/2025"
categories: []
image: ""
format: html
---

# Preamble {#sec-preamble}

## Install Libraries
```{r}
#install.packages("remotes")
#remotes::install_github("DevPsyLab/petersenlab")
#remotes::install_github("FantasyFootballAnalytics/ffanalytics")

#update.packages(ask = FALSE)
```

## Load Libraries
```{r}
#| message: false
#| warning: false

library("petersenlab")
library("tidyverse")
```

## Load Data

```{r}
#| eval: false

load(file = "../../data/player_stats_weekly.RData")
load(file = "../../data/players_projections_weekly.RData")
load(file = "../../data/nfl_playerIDs.RData")
```

## Prepare Data

```{r}
player_stats_weekly_subset <- player_stats_weekly %>% 
  filter(!is.na(player_id))

nfl_playerIDs_subset <- nfl_playerIDs %>% 
  filter(!is.na(gsis_id)) %>% 
  distinct(gsis_id, .keep_all = TRUE) %>% # keep only rows that do not have duplicate values of gsis_id
  select(-all_of(c("team", "position", "height", "weight", "age")))

players_projectedPoints_weekly_combined$season <- as.integer(players_projectedPoints_weekly_combined$season)
players_projections_weekly_average_merged$season <- as.integer(players_projections_weekly_average_merged$season)

player_stats_weekly_subset_IDs <- left_join(
  player_stats_weekly_subset,
  nfl_playerIDs_subset,
  by = c("player_id" = "gsis_id")
) %>% 
  filter(!is.na(mfl_id))

projectionsWithActuals_weekly <- full_join(
  player_stats_weekly_subset_IDs,
  players_projectedPoints_weekly_combined,
  by = c("mfl_id" = "id", "season", "week"),
  suffix = c("", "_proj"),
)

crowdAveragedProjectionsWithActuals_weekly <- full_join(
  player_stats_weekly_subset_IDs,
  players_projections_weekly_average_merged,
  by = c("mfl_id" = "id", "season", "week"),
  suffix = c("", "_proj"),
)

projectionsWithActuals_weekly <- projectionsWithActuals_weekly %>% 
  unite(
    "player_id_season_week",
    player_id,
    season,
    week,
    remove = FALSE
  )

crowdAveragedProjectionsWithActuals_weekly <- crowdAveragedProjectionsWithActuals_weekly %>% 
  unite(
    "player_id_season_week",
    player_id,
    season,
    week,
    remove = FALSE
  )
```

# A.I. Disclosure {#sec-aiDisclosure}

AI Disclosure: AI was not used.

Link to AI Transcript: not applicable


# Abstract {#sec-abstract}

**Background:**

**Method:**

**Results:**

**Discussion:**

# Background {#sec-background}

I want to use predictions to see the comparison, for wide receivers, between receptions, using the magic number of 5 from their last game, seeing if they are consistent on that number. Also the number of receiving yards, magic number of 60, and yards after catch, with 15. This is a thought-out comparison to see, if accuracy predictions, which one is better than the other, and can we safely say what is better to pick for a fantasy team.

## Research Question

Among NFL wide receivers, which prior-game signal—receptions ≥5, receiving yards ≥60, or yards after catch ≥15 most accurately and calibratedly predicts hitting the same threshold next game, and which signal yields the best start/sit decisions under binary-classification metrics

## Hypothesis

Prior-game receptions ≥5 will be the most accurate and best-calibrated predictor of repeating the threshold next game. 

## Justification

Leagues are in PPR (Points Per Reception) leagues, so the value of receptions is really high. Having a great amount of receptions per game.

# Method {#sec-method}



# Data {#sec-data}

```{r}
petersenlab::accuracyAtCutoff(
  predicted = player_stats_weekly_example$receptions_priorGame,
  actual = player_stats_weekly_example$rec_6,
  cutoff = 6
```

# Results {#sec-results}

# Discussion {#sec-discussion}

# Strengths and Limitations {#sec-strengthsAndLimitations}


# Citation {#sec-Citation}



# Session Info {#sec-sessionInfo}
```{r}
sessionInfo()
```
