---
title: "INSERT"
author: "Jordan Wozniak"
date: "10/30/2025"
categories: [INSERT]
image: ""
format: html
---

# Preamble {#sec-preamble}

## Install Libraries
```{r}
#install.packages("remotes")
#remotes::install_github("DevPsyLab/petersenlab")
#remotes::install_github("FantasyFootballAnalytics/ffanalytics")

#update.packages(ask = FALSE)
```

## Load Libraries
```{r}
#| message: false
#| warning: false

library("petersenlab")
library("tidyverse")
```

## Load Data

```{r}
load(file = "../../data/player_stats_weekly.RData")
load(file = "../../data/players_projections_weekly.RData")
load(file = "../../data/nfl_playerIDs.RData")
```

```{r}
options(scipen = 999) # prevent scientific notation
```

# A.I. Disclosure {#sec-aiDisclosure}

AI Disclosure: AI was not used.

Link to AI Transcript: not applicable

# Absract {#sec-abstract}

**Background**:
**Method**:
**Results**:
**Discussion**:

# Background {#sec-background}

## Research Question

## Hypothesis

## Justification

# Method {#sec-method}

# Results {#sec-results}

# Discussion {#sec-discussion}

# Strengths and Limitations {#sec-strengthsAndLimitations}

# Conclusion {#sec-conclusion}

# Citation {#sec-citation}

# Session Info {#sec-sessionInfo}